\pkg{RProtoBuf}: Efficient Cross-Language Data Serialization in \proglang{R}

### 2014-11-19 papers/jss/article.Rnw
Address referee #1 feedback about section 4.
Answer many of the specific questions from the referee: Why use S4 and
not RC (answer: written before RC availalbe, acknowledge that RC would
be better). objects made mutable by usual functional copy on modify
semantics in mutation methods, avoid pseudo-method.

Fix typographical issue with subheadings.  Make it more concise with
\subsection* and combine a few.



git-svn-id: svn://svn.r-forge.r-project.org/svnroot/rprotobuf@900 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\textbf{Importing message descriptors from \code{.proto} files}

@@

\\

\textbf{Creating, accessing, and modifying messages.}

@@

%\subsection{Access and modify fields of a message}

@@

%\subsection{Display messages}

\\

\textbf{Printing, Reading, and Writing Messages}

@@

% TODO(mstokely): comment out and use only one style, not both per
% referee feedback.  Also avoid using the term 'pseudo-method' which
% is unclear.
%
%\code{read} can also be used as a pseudo-method of the descriptor
%object:
%
%<<>>=
%message <- tutorial.Person$read(tf1)
%con <- file(tf2, open = "rb")
%message <- tutorial.Person$read(con)
%close(con)
%message <- tutorial.Person$read(payload)
%@
%
%Here we read first from a file, then from a binary connection and lastly from
%a message payload.

\section{Under the hood: S4 classes, methods, and pseudo methods}
\label{sec:rprotobuf-classes}

The \pkg{RProtoBuf} package uses the S4 system to store
information about descriptors and messages.  Using the S4 system
allows the package to dispatch methods that are not
generic in the S3 sense, such as \code{new} and
\code{serialize}.
Table~\ref{class-summary-table} lists the six
primary Message and Descriptor classes in \pkg{RProtoBuf}.  Each \proglang{R} object
contains an external pointer to an object managed by the
\code{protobuf} \proglang{C++} library, and the \proglang{R} objects make calls into more
than 100 \proglang{C++} functions that provide the
glue code between the \proglang{R} language classes and the underlying \proglang{C++}
classes.

@@

The \pkg{RProtoBuf} package supports two forms for calling
functions with these S4 classes:
\begin{itemize}
\item The functional dispatch mechanism of the the form
  \verb|method(object, arguments)| (common to \proglang{R}), and
\item The traditional object oriented notation
  \verb|object$method(arguments)|.
\end{itemize}

Additionally, \pkg{RProtoBuf} supports tab completion for all
classes.  Completion possibilities include pseudo-method names for all
classes, plus \emph{dynamic dispatch} on names or types specific to a given
object.  This functionality is implemented with the
\code{.DollarNames} S3 generic function defined in the \pkg{utils}
package that is included with \proglang{R} \citep{r}.

@@

Descriptors describe the type of a Message.  This includes what fields
a message contains and what the types of those fields are.  Message
descriptors are represented in \proglang{R} by the \emph{Descriptor} S4
class. The class contains the slots \code{pointer} and
\code{type}.  Similarly to messages, the \verb|$| operator can be
used to retrieve descriptors that are contained in the descriptor, or
invoke pseudo-methods.

@@

The \verb|$| operator can be used to retrieve the value of enum
constants contained in the EnumDescriptor, or to invoke
pseudo-methods.

@@

The \verb|$| operator can be used to invoke pseudo-methods.

@@

The \verb|$| operator can be used to retrieve named fields defined in
the FileDescriptor, or to invoke pseudo-methods.

>>>>

\subsection*{Importing message descriptors from \code{.proto} files}

@@

% \\

\subsection*{Creating, accessing, and modifying messages.}

@@

% \subsection*{Access and modify fields of a message}

@@

\subsection*{Printing, reading, and writing Messages}

%\\

% \textbf{Printing, Reading, and Writing Messages}

@@

% TODO(mstokely): comment out and use only one style, not both per
% referee feedback.  Also avoid using the term 'pseudo-method' which
% is unclear.
%
%\code{read} can also be used as a method of the descriptor
%object:
%
%<<>>=
%message <- tutorial.Person$read(tf1)
%con <- file(tf2, open = "rb")
%message <- tutorial.Person$read(con)
%close(con)
%message <- tutorial.Person$read(payload)
%@
%
%Here we read first from a file, then from a binary connection and lastly from
%a message payload.

\section{Under the hood: S4 classes and methods}
\label{sec:rprotobuf-classes}

The \pkg{RProtoBuf} package uses the S4 system to store
information about descriptors and messages.
Each \proglang{R} object
contains an external pointer to an object managed by the
\code{protobuf} \proglang{C++} library, and the \proglang{R} objects make calls into more
than 100 \proglang{C++} functions that provide the
glue code between the \proglang{R} language classes and the underlying \proglang{C++}
classes.
S4 objects are immutable, and so the methods that modify field values of a message return a new copy of the object with R's usual functional copy on modify semantics\footnote{RProtoBuf was designed and implemented before Reference Classes were introduced to offer a new class system with mutable objects.  If RProtoBuf were
implemented today Reference Classes would almost certainly be a better
design choice than S4 classes.}.
Using the S4 system
allows the package to dispatch methods that are not
generic in the S3 sense, such as \code{new} and
\code{serialize}.

Since \pkg{RProtoBuf} users are most often switching between two or
more different languages as part of a larger data analysis pipeline,
both generic function and message passing OO style calling conventions
are supported:

\begin{itemize}
\item The functional dispatch mechanism of the the form
  \verb|method(object, arguments)| (common to \proglang{R}), and
\item The traditional object oriented notation
  \verb|object$method(arguments)|.
\end{itemize}

Additionally, \pkg{RProtoBuf} supports tab completion for all
classes.  Completion possibilities include method names for all
classes, plus \emph{dynamic dispatch} on names or types specific to a given
object.  This functionality is implemented with the
\code{.DollarNames} S3 generic function defined in the \pkg{utils}
package that is included with \proglang{R} \citep{r}.

Table~\ref{class-summary-table} lists the six primary Message and
Descriptor classes in \pkg{RProtoBuf}.
% Please see the package
%documentation for a complete description of the slots and methods for
%each class.

@@

@@

Descriptors describe the type of a Message.  This includes what fields
a message contains and what the types of those fields are.  Message
descriptors are represented in \proglang{R} by the \emph{Descriptor} S4
class. The class contains the slots \code{pointer} and
\code{type}.  Similarly to messages, the \verb|$| operator can be
used to retrieve descriptors that are contained in the descriptor, or
invoke methods.

@@

The \verb|$| operator can be used to retrieve the value of enum
constants contained in the EnumDescriptor, or to invoke
methods.

@@

The \verb|$| operator can be used to invoke methods.

@@

The \verb|$| operator can be used to retrieve named fields defined in
the FileDescriptor, or to invoke methods.
